Formalization, Mechanization and Automation of G\"{o}del's Proof of God's Existence\thanks{This work has been supported by the German Research Foundation under grant BE2501/9-1. Bruno ????}

### 2013-08-18 Abstract/GodProof-Abstract.tex
minor modifications in abstract

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}
\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this sources.}
\allowdisplaybreaks[1] 
\begin{align}
& \text{Any property necessarily implied by a positive property is positive.} \notag \\
& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
& \text{A property is positive if and only if its negation is not positive.} \notag \\
& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
& \text{The property of being God-like is positive.} \notag \\
& \quad pos\,god \tag{A3} \\
& \text{Positive properties are necessarily positive.} \notag \\
& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
& \text{Necessary existence is a positive property:} \notag \\
& \quad pos\,nec\_exists \tag{A5} \\
& \text{x is God-like if and only if x has every positive property:} \notag \\
& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2}
\end{align}

>>>>

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}
\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this sources.}
\allowdisplaybreaks[1] 
\begin{align}
& \text{Any property necessarily implied by a positive property is positive.} \notag \\
& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
& \text{A property is positive if and only if its negation is not positive.} \notag \\
& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
& \text{The property of being God-like is positive.} \notag \\
& \quad pos\,god \tag{A3} \\
& \text{Positive properties are necessarily positive.} \notag \\
& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
& \text{Necessary existence is a positive property.} \notag \\
& \quad pos\,nec\_exists \tag{A5} \\
%\end{align}
%\begin{align}
& \text{x is God-like if and only if x has every positive property.} \notag \\
& \quad god\,x := \forall P (pos\,P \Rightarrow P\,x) \tag{D1} \\ 
& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box \forall y (P\,y \Rightarrow Q\,y)) \tag{D2} \\
& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
& \quad nec\_exists\,x := \forall P (ess\,P\,x \Rightarrow \Box \exists y (P\,y)) \tag{D3} 
\end{align}
